---
title: "Prediction of normal cells vs. tumor cells from human colon"
author: "Jose Alquicira Hernandez"
date: "05/10/2017"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 4
    toc_float: yes
    collapsed: false
    code_folding: show
---

# Run example

```{bash run_example, eval=FALSE}
# Set scripts directory ----
script_dir=bin

# Script parameters

# Set extra parameters -----

positive_class=normal
negative_class=tumor

# Input files --------------

exp_data=results/2017-10-05_colon_cancer_pca/coltumor_exp_data_processed.RDS
red_data=results/2017-10-05_colon_cancer_pca/pca_all_nm_vs_all_tumor.RDS

# Output file directory ----

output_path=results/${PWD##*/}


# Main ---------------------

Rscript -e "library('here'); rmarkdown::render(here('${script_dir}', 'coltumor_pca_prediction.Rmd'), output_file = paste0('${positive_class}', '-', '${negative_class}', '.html'), output_dir = here('${output_path}'))" ${red_data} ${exp_data} ${output_path} ${positive_class} ${negative_class}
```

```{r import_primary_libraries}
library("scPrediction")
library("tidyverse")
library("here")
library("stringr")
```

```{r command_line_args}
argv <- commandArgs(TRUE)

# Set input parameters

red.data <- argv[1]
exp.data <- argv[2]

# Set output parameters

output.path<- argv[3]

# Set extra parameters

positive.class <- argv[4]
negative.class <- argv[5]

# positive.class <- "normal"
# negative.class <- "tumor"
```

# Groups prediction {.tabset}

## Predict groups using eigenvectors

### Read and process input data

```{r read_pca_data}
red <- readRDS(here(red.data))
eigenvectors <- red$x
groups <- red$status
```

### Get significant eigenvectors

```{r get_significant_eigen}
red.features <- getSignificantEigen(dim.red = red, groups = groups)
```

### Make predictions

```{r make_predictions_eigen, results='hide', warning=FALSE, message=FALSE, cache=TRUE}
performance <- predictEigen(eigenvectors = eigenvectors,
                            groups = groups,
                            positive.class = positive.class,
                            significant.eigenvectors = red.features, 
                            n.rep = 10,
                            n.cores = 10)
```

### Save results

```{r save_results}
cat("Saving prediction results...\n")

# Save performance
output.file.name <- paste0("performance_", positive.class, "_vs_", negative.class, ".RDS")
output.file.name.path <- file.path(here(output.path), output.file.name)

saveRDS(performance, file = output.file.name.path)

# Save table of significant eigenvectors
output.file.name <- paste0("pvals_significant-group_", positive.class, "_vs_", negative.class, ".txt")
output.file.name.path <- file.path(here(output.path), output.file.name)
write.table(red.features$significant.eigenvectors, 
            file = output.file.name.path,
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

# Save talbe of top N eigenvectors
output.file.name <- paste0("pvals_all-group_", positive.class, "_vs_", negative.class, ".txt")
output.file.name.path <- file.path(here(output.path), output.file.name)
write.table(red.features$top.n.eigenvectors,
            file = output.file.name.path,
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

cat("Results saved...\n")
```

### Plot metrics

```{r plot_metrics}
plots <- plotMetrics(performance, eigen.info =  red.features)
```



## Predict groups using DE genes


### Read GTEx dataset

```{r read_gtex_data}
exp.metadata <- readRDS(here(exp.data))
exp <- exp.metadata$exp.data
metadata <- exp.metadata$metadata
```


### Read DE genes information


```{r read_de_genes}
de.genes <- c("PLA2G2A", "AGR2", "ATP5A1", "UQCR10", "C15orf48", "PRAC1", "LGALS4", "CKB", "PIGR", "MT1E", "MT1G", "MT2A", "C10orf99", "FABP1", "MT1X", "FXYD3", "WFDC2", "NDUFB1", "DBI", "SERF2", "TMSB10", "ATP5I", "MYL6", "S100A10", "B2M", "UBA52", "RPS29", "COX6C", "COX7C", "USMG5", "RPS23", "ATP5G1", "LGALS3", "SNORD13", "SNORD118", "S100P", "S100A11", "HNRNPH1", "ANKRD10", "SAT1", "RP11-1143G9.4", "LYZ", "HSPH1", "HSP90AA1", "HSPA8")
```

```{r extract_de_genes}
exp %>% 
  row.names() %>% 
  str_split("_") %>% 
  lapply("[", 2) %>% 
  unlist() -> exp.data.gene.list

exp.deg <- exp[match(de.genes, exp.data.gene.list),]
```


### Dataset summary

Number of differentially expressed genes: `r nrow(exp.deg)`
Number of samples: `r ncol(exp.deg)`


```{r display_status_info}
metadata %>%
  group_by(status) %>%
  count()
```

### Make predictions

```{r make_predictions, results='hide', warning=FALSE, message=FALSE}
de.performance <- predictDEGenes(em = exp.deg,
                                 groups = groups,
                                 positive.class = positive.class,
                                 n.rep = 10,
                                 n.cores = 10)
```

```{r save_results_deg}
cat("Saving prediction results...\n")

# Save performance
output.file.name <- paste0("performance_deg_", positive.class, "_vs_", negative.class, ".RDS")
output.file.name.path <- file.path(here(output.path), output.file.name)

saveRDS(de.performance, file = output.file.name.path)
cat("Results saved...\n")
```

```{r format_results}
de.performance %>%
  group_by(Model) %>%
  summarise_at(.vars = c("Accuracy", "Sensitivity", "Specificity", "Kappa", "AUC"),
               .funs = funs(median)) %>%
  filter(Model == "svmPoly") %>%
  ungroup(n.ev) %>%
  as.data.frame()-> de.performance.median

de.metrics <- gather(data = de.performance.median, key = "Metric", value = "Value", 2:6)
```

# Plot results

```{r add_de_metrics_to_plot}
p.accuracy <- plots$accuracy + geom_hline(yintercept = de.performance.median$Accuracy)
p.kappa <- plots$kappa + geom_hline(yintercept = de.performance.median$Kappa)
p.auc <- plots$auc + geom_hline(yintercept = de.performance.median$AUC)
```

```{r make_plor_grid, fig.height=4, fig.width=12}
p <- plot_grid(p.accuracy, p.auc, p.kappa, nrow = 1)
p
```


```{r save_metrics_plot}
output.file.name <- paste0("metrics_", positive.class, "_vs_", negative.class, ".png")
output.file.name.path <- file.path(here(output.path), output.file.name)

ggsave(filename = output.file.name.path,
       plot = p,
       width = 20,
       height = 4.5,
       device = "png",
       dpi = 350)
```

# Session information

```{r display_session_info}
options(width = 120)
devtools::session_info()
```