---
title: "PCA of human colorectal tumors"
author: "Jose Alquicira Hernandez"
date: "05/10/2017"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 4
    toc_float: yes
    collapsed: false
    code_folding: show
---

# Run example

```{bash run_example, eval=FALSE}
# Set scripts directory ----
script_dir=bin

# Script parameters

# Input files --------------

tumor=data/2017-10-95_colon/GSE81861_CRC_tumor_epithelial_cells_FPKM.csv
normal=data/2017-10-95_colon/GSE81861_CRC_NM_epithelial_cells_FPKM.csv
# Output file directory ----

output_path=results/${PWD##*/}


# Main ---------------------

Rscript -e "library('here'); rmarkdown::render(here('${script_dir}', 'coltumor_processing.Rmd'), output_file = paste0('normal', '-', 'tumor', '.html'), output_dir = here('${output_path}'))" ${tumor} ${normal} ${output_path}
```

```{r import_libraries}
library(tidyverse)
library(here)
library(data.table)
library(Biobase)
library(stringr)
```

```{r command_line_args}
argv <- commandArgs(TRUE)

# Set input parameters

tumor.path <- argv[1]
normal.path <- argv[2]

# Set output parameters

output.path <- argv[3]
```

# Read data


```{r include=FALSE}
tumor.path="data/2017-10-95_colon/GSE81861_CRC_tumor_epithelial_cells_FPKM.csv"
normal.path="data/2017-10-95_colon/GSE81861_CRC_NM_epithelial_cells_FPKM.csv"
```


```{r read_data}
tumor <- fread(here(tumor.path),
               sep = ",",
               check.names = TRUE,
               data.table = FALSE)

row.names(tumor) <- tumor$V1
tumor$V1 <- NULL


normal <- fread(here(normal.path),
      sep = ",",
      check.names = TRUE,
      data.table = FALSE)


row.names(normal) <- normal$V1
normal$V1 <- NULL
```


```{r get_summary} 
list(tumor = tumor, normal = normal) %>% 
lapply(dim) %>% 
  as.data.frame(row.names = c("Genes", "Cells"))
```


Are loci in both datasets the same and ordered : `r all(row.names(tumor) == row.names(normal))`


```{r merge_datasets}
expData <- cbind(normal, tumor)
```

```{r create_metadata}
metadata <- data.frame(status = c(rep("normal", ncol(normal)), rep("tumor", ncol(tumor))), 
                       row.names = colnames(expData))
```

```{r get_extra_metadata}
getField <- function(i){
  metadata %>% 
  row.names() %>% 
  str_split("__") %>%
  lapply("[", i) %>% 
  unlist()
}

fields <- lapply(1:2, getField)
metadata$id <- fields[[1]]
metadata$cellType <- fields[[2]]
```


Are cell ids identical `r !any(duplicated(metadata$id))`
Are loci identical `r !any(duplicated(row.names(expData)))`


```{r shows_overall_summary}
metadata %>% 
  group_by(status, cellType) %>% 
  count()
```



```{r create_expressionSet}
description <- data.frame(labelDescription= c("Malignancy status",
                                              "Short cell id",
                                              "Cell type"),
                          row.names=c("status", "cellId", "cellType"))

phenoData <- new("AnnotatedDataFrame", data = metadata, varMetadata = description)

expSet <- ExpressionSet(assayData = as.matrix(expData), phenoData = phenoData)
```


# Data processing

## Filtering

### Remove regions with zero RPKM values across all cells


```{r remove_zero-count_genes}
genesZero <- esApply(expSet, 1, function(gene) all(gene == 0))
expSet <- expSet[!genesZero,]
```


**`r sum(genesZero)`** loci removed with all zero-values across all cells.



### Remove loci not expressed in at least 10% of all cells


```{r remove_non-shared_genes}
nonLowExpressed <- rowSums(exprs(expSet) > 0) >= round(ncol(expSet) * 0.1)

expSet <- expSet[nonLowExpressed,]
```

**`r sum(nonLowExpressed)`** loci kept with regions not expressed in at least 10% of cells


### Extract stem/TA cells

```{r extract_stem-ta_cells}
expSet <- expSet[,expSet$cellType == "stemTA"]
```


# Final data summary

Total number of regions/genes: **`r nrow(expSet)`**
Total number of cells: **`r ncol(expSet)`**

```{r final_summary}
expSet$status %>% 
  table() %>% 
  as.data.frame() %>% 
  set_names(c("Status", "Number of cells"))
```


```{r save_results}
saveRDS(expSet, 
        file = file.path(here(output.path), "coltumor_exp_data_processed.RDS"))
```

```{r display_session_info}
options(width = 120)
devtools::session_info()
```
