---
title: "Sample origin prediction using GTEx dataset"
author: "Jose Alquicira Hernandez"
date: "13/09/2017"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 4
    toc_float: yes
    collapsed: false
    code_folding: show
---

# Run example

```{bash run_example, eval=FALSE}
# Set scripts directory ----
script_dir=bin

# Script parameters

# Set extra parameters -----

sample_1=thyroid
sample_2=whole_blood

# Input files --------------

gtex_data=results/2017-09-21_gtex_cleaning_data/gtex_clean.RDS
red_data=results/2017-09-27_gtex_pca_samples/pca_${sample_1}_vs_${sample_2}.RDS
deg_data=results/2017-09-27_gtex_de_genes_samples/${sample_1}_vs_${sample_2}.csv

# Output file directory ----

output_dir=results/${PWD##*/}


# Main ---------------------

Rscript -e "library('here'); rmarkdown::render(here('${script_dir}', 'gtex_pca_prediction.Rmd'), output_file = paste0('${sample_1}', '-', '${sample_2}', '.html'), output_dir = here('${output_dir}'))" ${red_data} ${deg_data} ${gtex_data} ${output_dir} ${sample_1} ${sample_2}
```

```{r import_primary_libraries}
library("scPrediction")
library("tidyverse")
library("here")
```

```{r command_line_args}
argv <- commandArgs(TRUE)

# Set input parameters

red.data <- argv[1]
deg.data <- argv[2]
gtex.data <- argv[3]

# Set output parameters

output.path<- argv[4]

# Set extra parameters

positive.class <- argv[5]
negative.class <- argv[6]

# positive.class <- "skin_sun_exposed_lower_leg"
# negative.class <- "skin_not_sun_exposed_suprapubic"
```

# Groups prediction {.tabset}

## Predict groups using eigenvectors

### Read and process input data

```{r read_pca_data}
red <- readRDS(here(red.data))
eigenvectors <- red$x
groups <- red$sample
```

### Get significant eigenvectors

```{r get_significant_eigen}
red.features <- getSignificantEigen(dim.red = red, groups = groups)
```

### Make predictions

```{r make_predictions_eigen, results='hide', warning=FALSE, message=FALSE}
performance <- predictEigen(eigenvectors = eigenvectors,
                            groups = groups,
                            positive.class = positive.class,
                            significant.eigenvectors = red.features, 
                            n.rep = 10,
                            n.cores = 10)
```

### Save results

```{r save_results}
cat("Saving prediction results...\n")

# Save performance
output.file.name <- paste0("performance_", positive.class, "_vs_", negative.class, ".RDS")
output.file.name.path <- file.path(here(output.path), output.file.name)

saveRDS(performance, file = output.file.name.path)

# Save table of significant eigenvectors
output.file.name <- paste0("pvals_significant-group_", positive.class, "_vs_", negative.class, ".txt")
output.file.name.path <- file.path(here(output.path), output.file.name)
write.table(red.features$significant.eigenvectors, 
            file = output.file.name.path,
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

# Save talbe of top N eigenvectors
output.file.name <- paste0("pvals_all-group_", positive.class, "_vs_", negative.class, ".txt")
output.file.name.path <- file.path(here(output.path), output.file.name)
write.table(red.features$top.n.eigenvectors,
            file = output.file.name.path,
            sep = "\t",
            row.names = FALSE,
            quote = FALSE)

cat("Results saved...\n")
```

### Plot metrics

```{r plot_metrics}
plots <- plotMetrics(performance, eigen.info =  red.features)
```



## Predict groups using DE genes


### Read GTEx dataset

```{r read_gtex_data}
gtex <- readRDS(here(gtex.data))
```

```{r subset_data}
samples.idx <- (gtex$metadata$subtissue == positive.class) | (gtex$metadata$subtissue == negative.class)
metadata <- gtex$metadata[samples.idx, ]
tpm <- gtex$tpm[,samples.idx]
groups <- metadata$subtissue
```

### Read DE genes information


```{r read_de_genes}
de.genes <- read.csv(here(deg.data))
```


```{r extract_de_genes}
tpm.deg <- tpm[rownames(tpm) %in% de.genes$genes,]
```

### Dataset summary

Number of differentially expressed genes: `r nrow(tpm.deg)`
Number of samples: `r ncol(tpm.deg)`


```{r display_sample_info}
metadata %>%
  group_by(SMTSD) %>%
  count()
```

### Make predictions

```{r make_predictions, results='hide', warning=FALSE, message=FALSE}
de.performance <- predictDEGenes(em = tpm.deg,
                                 groups = groups,
                                 positive.class = positive.class,
                                 n.rep = 10,
                                 n.cores = 10)
```

```{r save_results_deg}
cat("Saving prediction results...\n")

# Save performance
output.file.name <- paste0("performance_deg_", positive.class, "_vs_", negative.class, ".RDS")
output.file.name.path <- file.path(here(output.path), output.file.name)

saveRDS(de.performance, file = output.file.name.path)
cat("Results saved...\n")
```

```{r format_results}
de.performance %>%
  group_by(Model) %>%
  summarise_at(.vars = c("Accuracy", "Sensitivity", "Specificity", "Kappa", "AUC"),
               .funs = funs(median)) %>%
  filter(Model == "svmPoly") %>%
  ungroup(n.ev) %>%
  as.data.frame()-> de.performance.median

de.metrics <- gather(data = de.performance.median, key = "Metric", value = "Value", 2:6)
```

# Plot results

```{r add_de_metrics_to_plot}
p.accuracy <- plots$accuracy + geom_hline(yintercept = de.performance.median$Accuracy)
p.kappa <- plots$kappa + geom_hline(yintercept = de.performance.median$Kappa)
p.auc <- plots$auc + geom_hline(yintercept = de.performance.median$AUC)
```

```{r make_plor_grid, fig.height=4, fig.width=12}
p <- plot_grid(p.accuracy, p.auc, p.kappa, nrow = 1)
p
```


```{r save_metrics_plot}
output.file.name <- paste0("metrics_", positive.class, "_vs_", negative.class, ".png")
output.file.name.path <- file.path(here(output.path), output.file.name)

ggsave(filename = output.file.name.path,
       plot = p,
       width = 20,
       height = 4.5,
       device = "png",
       dpi = 350)
```

# Session information

```{r display_session_info}
options(width = 120)
devtools::session_info()
```


