---
title: "Get DE genes from GTEx dataset"
author: "Jose Alquicira Hernandez"
date: "Last run: `r format(Sys.time(), '%b %d %Y')`"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 4
    toc_float: yes
    collapsed: false
    code_folding: show
---

# Run example

```{bash run_example, eval=FALSE}
# Set scripts directory ----
script_dir=bin

# Input files parameters ---

gtex_path=results/2017-09-21_gtex_cleaning_data/gtex_clean.RDS

# Output file parameters ---

output_dir=results/${PWD##*/}

# Set extra parameters -----

sample_1=brain_frontal_cortex_ba9 
sample_2=brain_cortex

# Main ---------------------

Rscript -e "library('here'); rmarkdown::render(here('${script_dir}', 'gtex_get_de_genes.Rmd'), output_file = paste0('${sample_1}', '-', '${sample_2}', '.html'), output_dir = here('${output_dir}'))" ${gtex_path} ${output_dir} ${sample_1} ${sample_2}
```



# Import libraries

```{r import_primary_libraries}
library("here")
library("tidyverse")
```



```{r import_secondary_libraries}
library("edgeR")
library("stringr")
library("knitr")
```

# Read command-line arguments

```{r command_line_args}
argv <- commandArgs(TRUE)

# Set input parameters
path.gtex.data <- argv[1]

# Set output parameters
output.path<- argv[2]

# Set extra parameters
positive.class <- argv[3]
negative.class <- argv[4]
# positive.class <- "brain_frontal_cortex_ba9"
# negative.class <- "brain_cortex"
```


# Analyzed groups

- **Positive class**: `r positive.class`
- **Negative class**: `r negative.class`


# Read data

```{r read_gtex_data}
gtex <- readRDS(here(path.gtex.data))
```

```{r subset_data}
samples.idx <- (gtex$metadata$subtissue == positive.class) | (gtex$metadata$subtissue == negative.class)
metadata <- gtex$metadata[samples.idx, ]
tpm <- gtex$tpm[,samples.idx]
raw.counts <- gtex$counts[,samples.idx]
```

```{r}
counts <- DGEList(raw.counts)
counts$samples$group <- as.factor(metadata$subtissue)
```



# Normalize counts


```{r calculate_norm_factors}
counts.norm <- calcNormFactors(counts, method = "TMM")
```

```{r create_design_matrix}
design <- model.matrix(~0 + group, data = counts.norm$samples)
colnames(design) <- levels(counts.norm$samples$group)
```


```{r voom}
voom.out <- voom(counts.norm, design, plot = TRUE)
```

# Get differentially expressed genes

```{r get_DE_genes}
fit <- lmFit(voom.out)

cont.matrix <- makeContrasts(contrasts = paste0(positive.class, " - ", negative.class), 
                             levels = design)

fit.cont <- contrasts.fit(fit, cont.matrix)
fit.cont <- eBayes(fit.cont)

summa.fit <- decideTests(fit.cont)
summary(summa.fit)
```


```{r plot_final_model}
plotSA(fit.cont, main="Final model: Mean−variance trend")
```


```{r get_results}
limma.res <- topTable(fit.cont, coef = 1, sort.by = "p", n = "Inf")
```


```{r show_results_top10}
topTable(fit.cont, coef = 1, sort.by = "p", n = 10)
```

```{r plot_DE_genes}
plotMD(fit.cont, coef = 1, status = summa.fit[,1], main = paste(positive.class, "vs.", negative.class))
```

```{r volcano_plot}
volcanoplot(fit.cont, coef = 1, highlight = 100, 
            names = unlist(lapply(strsplit(rownames(fit.cont$coefficients), "_"), "[", 2)),
            main = paste(positive.class, "vs.", negative.class))
```

```{r get_significant_genes}
#get significant DE genes
limma.res %>% 
  mutate(genes = rownames(.)) %>% 
  filter(adj.P.Val < 0.05, abs(logFC) > 2) -> significant.genes
```

```{r save_results}
output.file.name <- paste0(positive.class, "_vs_", negative.class, ".csv")
output.file.name.path <- file.path(here(output.path), output.file.name)
write.csv(significant.genes, file = output.file.name.path, row.names = FALSE)
```


# Number of significant genes

After filtering by adjusted p-value using a threshold of `0.05` and an absolute fold-change of `2` the number of differentially expressed genes between `r positive.class` and `r negative.class` is:

**`r nrow(significant.genes)`**

# Session info

```{r display_session_info}
options(width = 120)
devtools::session_info()
```

