---
title: "Plot prediction results of HBCA"
author: "Jose Alquicira Hernandez"
date: "27/11/2017"
output: html_document
---


```{r import_libraries, message=FALSE}
library("tidyverse")
library("here")
library("scPrediction")
library("cowplot")
library("knitr")
```


# Read data

```{r set_cell_types}
cellTypes <- c(paste0("DC", seq_len(6)), paste0("Mono", seq_len(4)))
```


## Eigenpredictions

```{r read_eigenPred}
readEigenPred <- function(cellType){
  eigenPred <- readRDS(here(paste0("results/2017-11-26_hbca_prediction/performance_",
                                   cellType,
                                   "_vs_other.RDS")))
  eigenPred
}

eigenPred <- lapply(cellTypes, readEigenPred)
names(eigenPred) <- cellTypes
```


## DEG predictions

```{r read_degPred}
readDEGPred <- function(cellType){
  degPred <- readRDS(here(paste0("results/2017-11-26_hbca_prediction/performance_DEG_", 
                               cellType,
                               "_vs_other.RDS")))
}

degPred <- lapply(cellTypes, readDEGPred)
names(degPred) <- cellTypes
```


# Plot data


```{r plotMetrics, eval=FALSE}
plotPerformance <- function(eigen, DEG, n){
  
  accPlot <- plotMetrics(pred = eigen, predDEG = DEG, metric = "accuracy")
  rocPlot <- plotMetrics(pred = eigen, predDEG = DEG,  metric = "roc")
  diagnosticsPlots <- plotDiagnostics(pred = eigen,  metric = "roc")
  
  
  list(accuracy = accPlot,
       roc = rocPlot,
       deltaCumVarPval = diagnosticsPlots$plotVarPval,
       deltaPerformance = diagnosticsPlots$plotDeltaPerformance)
}



performancePlots <- mapply(plotPerformance,
                      eigenPred,
                      degPred,
                      SIMPLIFY = FALSE)

```



Based on the diagnostic plots, a number of eigenvectors were selected for each cell type prediction:

| Cell type | Number of eigenvectors |
|-----------|------------------------|
| DC1       | 6                      |
| DC2       | 8                      |
| DC3       | 6                      |
| DC4       | 8                      |
| DC5       | 7                     |
| DC6       | 4                      |
| Mono1     | 6                      |
| Mono2     | 10                     |
| Mono3     | 7                      |
| Mono4     | 6                      |



```{r}
numEigen <- c(6, 8, 6, 8, 7, 4, 6, 10, 7, 6) 
names(numEigen) <- cellTypes
```

```{r plotMetrics}
plotPerformance <- function(eigen, DEG, n){
  
  #accPlot <- plotMetrics(pred = eigen, predDEG = DEG, n = n, metric = "accuracy")
  rocPlot <- plotMetrics(pred = eigen, predDEG = DEG,  n = n, metric = "roc")
  diagnosticsPlots <- plotDiagnostics(pred = eigen, n = n, metric = "roc")
  
  
  list(roc = rocPlot,
       deltaCumVarPval = diagnosticsPlots$plotVarPval,
       deltaPerformance = diagnosticsPlots$plotDeltaPerformance)
}



performancePlots <- mapply(plotPerformance,
                      eigenPred,
                      degPred,
                      numEigen,
                      SIMPLIFY = FALSE)

```



```{r}
cellType <- performancePlots$DC1
cellType$deltaCumVarPval$theme$axis.title.y.right <- element_blank()

DC1Plot <- plot_grid(cellType$roc + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)),
          cellType$deltaCumVarPval + xlab("") + ylab("") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)), 
          cellType$deltaPerformance + xlab("") + ylab("") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)),
          nrow = 1,
          align = "v", labels = LETTERS[1:3])

cellType <- performancePlots$Mono4
cellType$deltaCumVarPval$theme$axis.title.y.right <- element_blank()


Mono4Plot <- plot_grid(cellType$roc + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)),
          cellType$deltaCumVarPval + xlab("") + ylab("") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)), 
          cellType$deltaPerformance + xlab("") + ylab("") +  theme(axis.text.x = element_text(angle = 90, hjust = 1)),
          nrow = 1,
          align = "v")


plotClassRes <- function(cellType){
  cellType$deltaCumVarPval$theme$axis.title.y.right <- element_blank()
  plot_grid(cellType$roc + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)),
            cellType$deltaCumVarPval + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)), 
            cellType$deltaPerformance + xlab("") + ylab("") + theme(axis.text.x = element_text(angle = 90, hjust = 1)),
            nrow = 1,
            align = "v")
}


DCPlots <- lapply(performancePlots[2:9], plotClassRes)

DCPlots$DC1 <- DC1Plot
DCPlots$Mono4 <- Mono4Plot

tmp <- DCPlots[sort(names(DCPlots))]



resPlot <- plot_grid(plotlist = tmp, ncol = 1)
```


```{r}
ggsave(filename = here(paste0("results/2017-11-26_hbca_prediction/",
                              "performance_all_cell_types", ".png")),
       plot = resPlot, width = 10, height = 13, dpi = 350)
```

# Extract results

## Selected PCs

```{r}
getROC <- function(cellType, numEigen){
  lapply(cellType, function(x){
    x$predictions[[numEigen]]$roc.auc 
  }) %>% unlist()
}

rocNumEigen <- mapply(getROC, eigenPred, as.character(numEigen), SIMPLIFY = FALSE)
```

## DE genes

```{r}
getAUCDEG <- function(x){
  lapply(x, function(x) x["roc.auc"]) %>% unlist()
}

degPred %>% 
  lapply(getAUCDEG)  -> performanceDEG
```


## Exploratory plot


Plot *area under the ROC curve* distributions stratified by cell types and methods (eigen predictions vs. DEG-based predictions)

```{r}
library(ggridges)
library(RColorBrewer)

colPalette <- brewer.pal(12, name = "Set3")
colPalette <- colPalette[5:6]

rocNumEigen <- as.data.frame(rocNumEigen)
performanceDEG <- as.data.frame(performanceDEG)



rocNumEigen %>% 
  gather(key = "cellType", value = "roc") %>% 
  mutate(method = "PCs") -> rocNumEigenLong

performanceDEG %>% 
  gather(key = "cellType", value = "roc") %>% 
  mutate(method = "DEGs") -> performanceDEGLong


performanceAll <- rbind(rocNumEigenLong, performanceDEGLong)

performanceAll %>% 
  mutate(method = factor(method, levels = rev(unique(method))),
         cellType = factor(cellType, levels = rev(unique(cellType)))) -> performanceAll


p <- ggplot(performanceAll, aes(x = roc, y = cellType, fill = method)) +
  geom_density_ridges(alpha = 0.5) +
  scale_fill_manual(values  = colPalette, breaks = c("PCs","DEGs")) +
  xlab("AUC ROC") +
  ylab("Cell type") +
  theme_bw()



ggsave(filename = here(paste0("results/2017-11-26_hbca_prediction/",
                              "predictions_comparisons", ".png")),
       plot = p, width = 9, height = 8,
       dpi = 350)
```





# Test for distribution differences

```{r}
testDiff <- function(eigenRes, degRes){
  testRest <- wilcox.test(eigenRes, degRes[!is.na(degRes)], alternative = "greater")
}


diffs <- mapply(testDiff, rocNumEigen, performanceDEG, SIMPLIFY = FALSE)

diffs %>% 
  lapply("[[", "p.value") %>% 
  as.vector() %>% 
  p.adjust(method =  "bonferroni") %>% 
  as.data.frame() %>% 
  set_names(c("pValues")) %>% 
  mutate(cellType = row.names(.)) %>% 
  .[,c(2,1)] %>% 
  .["pValues"] %>% 
  format(digits = "3") %>% 
  as.vector()
```

