---
title: "Data processing of human monocytes and dendritic cells"
author: "Jose Alquicira Hernandez"
date: "24/11/2017"
output:
  html_document:
    df_print: kable
    fig_caption: yes
    highlight: kate
    number_sections: yes
    theme: simplex
    toc: yes
    toc_depth: 4
    toc_float: yes
    collapsed: false
    code_folding: show
---

# Run example

```{bash run_example, eval=FALSE}
# Set scripts directory ----
scriptDir=bin

# Script parameters

# Input files --------------

expDataDir=data/2017-07-17_blood_atlas/RAW/raw_expression_matrix.txt
metaDataDir=data/2017-07-17_blood_atlas/RAW/metadata.txt

# Output file directory ----

outputDir=results/${PWD##*/}


# Main ---------------------

Rscript -e "library('here'); rmarkdown::render(here('${scriptDir}', 'hbca_data_processing.Rmd'), output_file = paste0('blood_atlas', '.html'), output_dir = here('${outputDir}'))" ${expDataDir} ${metaDataDir} ${outputDir}
```

```{r import_libraries}
library("tidyverse")
library("here")
library("data.table")
library("Biobase")
library("stringr")
```

```{r command_line_args}
argv <- commandArgs(TRUE)

# Set input parameters

expDataDir<- argv[1]
metaDataDir <- argv[2]

# Set output parameters

outputDir<- argv[3]
```

# Read data

```{r read_data}
expDataMatrix <- read.table(here(expDataDir),
                 header = TRUE,
                 sep = "\t")


metadata <- read.table(here(metaDataDir),
                       header = TRUE,
                       sep = "\t",
                       skip = 1)

rownames(expDataMatrix) <- make.names(expDataMatrix$Gene.ID, unique = TRUE)
expDataMatrix$Gene.ID <- NULL
```



```{r filter_cells_by_metadata}
cellNames <- names(expDataMatrix) %>% 
  str_split("_") %>% lapply(function(x) paste0(x[seq_len(length(x) - 1)], collapse = "_")) %>% 
  unlist()

names(expDataMatrix) <- cellNames
expDataMatrixFilter <- expDataMatrix[, names(expDataMatrix) %in% metadata$TYPE]

dim(expDataMatrix)
dim(expDataMatrixFilter)
```

`r table(cellNames %in% metadata$TYPE)[1]` cells have no associated metadata. These cells were removed.
Do all cells from the metadata intersect with the cells in the gene expression matrix? : `r all(metadata$TYPE %in% cellNames)`

Order metadata according to gene expression matrix

```{r order_metadata}
metadata <- metadata[match(names(expDataMatrixFilter), metadata$TYPE), ]
rownames(metadata) <- metadata$TYPE
metadata$TYPE <- NULL
```
Are all cells from the metadata in the same order as in the gene expression matrix? : `r all(metadata$TYPE == names(expDataMatrixFilter))`



- Are cell ids identical `r !any(duplicated(metadata$TYPE))`
- Are loci identical `r !any(duplicated(rownames(expDataMatrixFilter)))`


```{r shows_overall_summary}
metadata %>% 
  group_by(group) %>% 
  count(sort = TRUE)
```



```{r create_expressionSet}
description <- data.frame(labelDescription= c("Cell type"),
                          row.names=c("cellType"))
names(metadata) <- "cellType"

phenoData <- new("AnnotatedDataFrame", data = metadata, varMetadata = description)

expDataMatrixFilter <- as.matrix(round(expDataMatrixFilter))
expSet <- ExpressionSet(assayData = expDataMatrixFilter, phenoData = phenoData)
```


# Data processing

## Filtering

### Remove regions with zero counts across all cells


```{r remove_zero-count_genes}
genesZero <- esApply(expSet, 1, function(gene) all(gene == 0))
expSet <- expSet[!genesZero,]
```


**`r sum(genesZero)`** loci removed with all zero-values across all cells.



### Remove loci not expressed in at least 10% of all cells


```{r remove_non-shared_genes}
nonLowExpressed <- rowSums(exprs(expSet) > 0) >= round(ncol(expSet) * 0.1)

expSet <- expSet[nonLowExpressed,]
```

**`r sum(nonLowExpressed)`** loci kept with regions not expressed in at least 10% of cells


# Final data summary

- Total number of regions/genes: **`r nrow(expSet)`**
- Total number of cells: **`r ncol(expSet)`**

```{r final_summary}
expSet$cellType %>% 
  table() %>% 
  as.data.frame() %>% 
  set_names(c("Cell type", "Number of cells"))
```


```{r create_cpm_expressionSet}
cpm  <- esApply(expSet, 2, function(x) (x/sum(x))*1000000)
expSetCPM <- expSet
exprs(expSetCPM) <- cpm
```


```{r save_results_counts}
saveRDS(expSet, 
        file = file.path(here(outputDir), "hbca_exp_data_processed_counts.RDS"))
```


```{r save_results_cpm}
saveRDS(expSetCPM, 
        file = file.path(here(outputDir), "hbca_exp_data_processed_cpm.RDS"))
```

```{r display_session_info}
options(width = 120)
devtools::session_info()
```
