---
title: "Plot prediction results of human colorectal cancer cells"
author: "Jose Alquicira Hernandez"
date: "15/01/2018"
output: html_document
---


```{r import_libraries, message=FALSE}
library("tidyverse")
library("here")
library("scPrediction")
library("cowplot")
library("knitr")
```


# Read data


## Eigenpredictions

```{r read_eigenPred}
eigenPred <- readRDS(here("results/2017-11-22_colon_cancer_prediction/performance_tumor_vs_normal.RDS"))
```


## DEG predictions

```{r read_degPred}
degPred <- readRDS(here(paste0("results/2018-03-13_hcc_DEGs_pred/performance_DEG_tumor_vs_normal.RDS")))
```


# Plot data



```{r plotMetrics, eval=FALSE}
accPlot <- plotMetrics(pred = eigenPred, predDEG = degPred, metric = "accuracy")
rocPlot <- plotMetrics(pred = eigenPred, predDEG = degPred,  metric = "roc")
diagnosticsPlots <- plotDiagnostics(pred = eigenPred,  metric = "roc")
```


```{r}
accPlot
rocPlot
diagnosticsPlots$plotVarPval
diagnosticsPlots$plotDeltaPerformance
```




Based on the diagnostic plots, the number of eigenvectors selected was `13`.


```{r plotMetrics}
n <- 13
rocPlot <- plotMetrics(pred = eigenPred, predDEG = degPred,  n = n, metric = "roc")
diagnosticsPlots <- plotDiagnostics(pred = eigenPred, n = n, metric = "roc")
```



```{r}
cancerControl <- plot_grid(diagnosticsPlots$plotVarPval + xlab("") + 
                             theme(axis.text.x = element_text(angle = 90, hjust = 1)),
                           rocPlot + xlab("") +  ylab("AUROC") +
                             theme(axis.text.x = element_text(angle = 90, hjust = 1)),
                           diagnosticsPlots$plotDeltaPerformance + ylab("Delta AUROC") +
                             theme(axis.text.x = element_text(angle = 90, hjust = 1)),
                           ncol = 1,
                           align = "v",
                           labels = LETTERS[1:3])
```


```{r}
ggsave(filename = here(paste0("results/2017-11-22_colon_cancer_prediction/",
                              "performance_tumor_vs_normal", ".png")),
       plot = cancerControl, width = 6, height = 8, dpi = 350)
```

# Extract results

## Selected PCs

```{r}
rocNumEigen <-  lapply(eigenPred, function(x){
  x$predictions[[n]]$roc.auc 
}) %>% unlist()
```


```{r}
 lapply(eigenPred, function(x){
  x$expTrainRedSig$topEigenvectors$EV}) %>%
  unlist() %>%
  table() %>%
  as.data.frame() %>%
  set_names(c("PC", "Freq")) -> pcInfo
  
  pcInfo$PC <- factor(pcInfo$PC, levels = as.character(pcInfo[order(pcInfo$Freq, decreasing = TRUE), "PC"]))

ggplot(pcInfo, aes(PC, Freq)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Frequency of PCs selection for human colorectal cancer along 50 replicates", subtitle = "13 principal components")

```

## Get best model


```{r}
eigenPred[[which.max(rocNumEigen)]]$expTrainRedSig$topEigenvectors$variance %>% 
  sum()
```


## DE genes

```{r}
degPred %>% 
  lapply(function(x) x["roc.auc"]) %>%
  unlist()  -> performanceDEG
```


## Exploratory plot


Plot *area under the ROC curve* distributions stratified by cell types and methods (eigen predictions vs. DEG-based predictions)

```{r}
library(ggridges)
library(RColorBrewer)

colPalette <- brewer.pal(12, name = "Set3")
colPalette <- colPalette[5:6]




performanceAll <- data.frame(eigen = rocNumEigen, deg = performanceDEG)

performanceAll %>% 
  gather(key = "method", value = "Roc") %>% 
  mutate(method = factor(method, levels = rev(unique(method)))) -> performanceAll


p <- ggplot(performanceAll, aes(x = Roc, fill = method)) +
  geom_density(alpha = 0.5) +
  scale_fill_manual(values  = colPalette, breaks = c("eigen","deg"), labels = c("PCs", "DEGs")) +
  xlab("AUC ROC") +
  ylab("Density") +
  theme_bw()


ggsave(filename = here(paste0("results/2017-11-22_colon_cancer_prediction/",
                              "predictions_comparisons", ".png")),
       plot = p, width = 6, height = 5,
       dpi = 350)
```



## Paper figures

```{r}
performanceAll %>% 
  mutate(method = factor(method, levels = c("eigen","deg"))) %>% 
  ggplot(aes(x = method, y = Roc)) +
  geom_boxplot() +
  geom_jitter() + 
  scale_x_discrete(labels = c("scPred", "DEGs")) +
  xlab("") +
  ylab("") -> p

ggsave(filename = here(paste0("results/2017-11-22_colon_cancer_prediction/",
                              "results_hcc", ".png")),
       plot = p, width = 5, height = 6,
       dpi = 350)

```






# Test for distribution differences

```{r}
testRest <- wilcox.test(rocNumEigen, performanceDEG[!is.na(performanceDEG)], alternative = "greater")

testRest$p.value
```
# Get median differences between groups

```{r}
median(rocNumEigen) - median(performanceDEG, na.rm = TRUE)
```
