---
title: "prediction"
author: "Jose Alquicira Hernandez"
date: "11 May 2017"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
# require("knitr")
# opts_knit$set(root.dir = "/Users/j.alquicira/Documents/powell_lab/projects/cell_prediction")
```


```{r load_libraries}
library('glmnet')
library('foreach')
library('doParallel')
require('dplyr')
```


```{r command_line_args}
argv <- commandArgs(TRUE)
cluster.ID <- argv[1]
day.ID <- argv[2]
```


```{r set_variables}
# Set variables
# cluster.ID <- 1
# day.ID <- 2
```


# Read data

- `expression.data` corresponds to a gene expression matrix. Rows are *genes* and columns are *cells*
- `cell.clusters` is a vector. Each value corresponds to one cluster. The length of the vector is the total of cell


```{r read_data}
# expression.data <- readRDS('ori_dat_sample.RDS')
# cell.clusters <- readRDS('my.clusters_0.45_Day2.RDS')
expression.data <- readRDS(paste0("Exprs_DCVLnorm_unlog_minus1_pos_Day", day.ID, ".RDS"))
cell.clusters <- readRDS(paste0("my.clusters_0.45_Day",day.ID, ".RDS"))
```

```{r show_unique_clusters}
cat("Clusters:\n")
cat(sort(unique(cell.clusters)))
cat("\n")
```

```{r format_gene_names}
rownames(expression.data) <- gsub("_.*", '', rownames(expression.data))
```


```{r get_cluster_ids}
n.clusters <- length(unique(cell.clusters))
```


```{r define_clusters}
## `cluster_select` contains indexes of all cluster tags that are equal to the cluster to be evaluated
cluster.select <- which(cell.clusters == as.numeric(cluster.ID))

## `cluster_compare` contains indexes of remaining cluster tags to be compared
cluster.compare <- which(cell.clusters != as.numeric(cluster.ID))
```


```{r cell-cluster_classification}
## cell.ids.cluster is a dataframe with two columns:
### 1. Cell id
### 2. Assigned cluster
cell.ids.cluster <- cbind(colnames(expression.data), cell.clusters)
```


```{r read_DEdata}
# Read differentially expressed data
DE.genes <- read.table(paste0('DEseq_Cluster', cluster.ID, '_vs_OtherClusters_Day', day.ID,'.txt_filtered_pAdjusted_sorted.txt'), header = TRUE)
# DE.genes <- readRDS('DE_result_sample.RDS')

# Gets gene names from `DE_result` and gets indexes of differential expressed genes in `expression.data` 
DE.gene.names <- gsub("_.*", '',  DE.genes$id)
DE.gene.idx <- which(rownames(expression.data) %in% DE.gene.names)
```



# Run regularization method

```{r init_params}
# Assign values to variables
clusterID.select <- cluster.ID
clusterID.comp <- '2'
features <- DE.gene.idx
```


# `BuildPredictor` function

## Parameters

- `clusterID.select`: ID of cluster of interest
- `cluster.select`: Indexes of cells from cluster of interest
- `clusterID.comp`: ID of cluster to be compared
- `cluster.compare`: Indexes of cells from cluster of interest
- `features`: Indexes for features to be included in the model (list of genes of interes, differentially expressed genes)
- `expression.data`: Gene expression matrix
  + Rows: genes
  + Columns: cells

## Output


```{r define_BuildTrainTest}

BuildTrainTest <- function(clusterID.select, cluster.select, clusterID.comp, cluster.compare, features, expression.data){
  
  # Random sampling for training data --------------------------------------
  # Take random sample of cells corresponding to cluster of interest
  
  # Set seed manually
  seed <- sample(1:2^15, 1)
  cat(sprintf("Seed for cluster select: %s\n", seed))
  set.seed(seed)

  cluster.select.indx <- sample(cluster.select, 
                                size = round(length(cluster.select)/2), 
                                replace = FALSE)
  # Set seed manually
  seed <- sample(1:2^15, 1)
  cat(sprintf("Seed for cluster compare: %s\n", seed))
  set.seed(seed)
  
  
  # Take random sample of cells corresponding to cluster to be compared
  cluster.compare.indx <- sample(cluster.compare, 
                                 size = round(length(cluster.compare)/2), 
                                 replace = FALSE)
  
  # Build predictor matrix -------------------------------------------------
  # Prepare predictor matrix containing both clutering classes
  predictor <- expression.data[features, c(cluster.select.indx, cluster.compare.indx)]
  
  # Generate categorical response ------------------------------------------
  # Set all values to cluster of interest
  response <- rep(clusterID.select, ncol(predictor))
  cluster.compare.names <- colnames(expression.data[,cluster.compare])
  sub.clustercompare.indx <- which(colnames(predictor) %in% cluster.compare.names)
  
  # Replace values for cluster.compare
  response[sub.clustercompare.indx] <- rep(clusterID.comp, length(sub.clustercompare.indx))

  # tranpose prediction matrix
  predictor <- t(predictor)
  
  # Random sampling for testing data ---------------------------------------
  # Prepare validation test. keep all cells except for those used in the training set
  cluster.select.indx.val <- sample(cluster.select[-cluster.select.indx])
  cluster.compare.indx.val <- sample(cluster.compare[-cluster.compare.indx])
  test <- expression.data[features, c(cluster.select.indx.val , cluster.compare.indx.val)]
  test <- t(test)
  
  # Return results ---------------------------------------------------------
  return(list(training = predictor, test = test, response = response))
}
```


# `FitRegModel` function

## Parameters

- `predictor`: gene expression matrix used as predictor
- `response`: vector of categorical values
- `clusterID.select`: ID of cluster of interest
- `cluster.select`: Indexes of cells from cluster of interest
- `clusterID.comp`: ID of cluster to be compared
- `cluster.compare`: Indexes of cells from cluster of interest



```{r}
FitRegModel <- function(test, training, response, ...){
  
  # Fit model
  fit <- glmnet(x = training, 
                y = response, 
                ...)
  
  # Fitting with cross validation
  cvfit <- cv.glmnet(x = training, 
                     y = response, 
                     type.measure = "class", ...)
  
  # Extract Beta coefficients using an optimized lambda value 
  cvfit.out <- as.data.frame(as.matrix(coef(cvfit, s = cvfit$lambda.min)))
  
  #find genes with coefficient different to 0
  cvfit.out$name <- row.names(cvfit.out)
  sub.cvfit.out <- cvfit.out[cvfit.out$`1` != 0,]
  
  # Extract explained deviance
  temp <- capture.output(cvfit.dev <- as.data.frame(as.matrix(print(cvfit$glmnet.fit))))
  names(cvfit.dev) <- c('df', 'deviance', 'lambda')
  
  # Get the coordinate for lambda that produces minimum error 
  cvfit.dev.lambda.idx <- which(round(cvfit.dev$lambda,digit = 3) == round(cvfit$lambda.min,digits = 3))
  cvfit.dev <- cvfit.dev[1:cvfit.dev.lambda.idx[1],]
  
  
  # Get max deviance for each degree of freedom level
  suppressMessages(require(dplyr))
  
  cvfit.dev %>% 
    group_by(df) %>% 
    summarise(deviance = max(deviance)) %>% 
    as.data.frame() -> cvfit.dev.max
  
  
  cvfit.dev.max$DEgenes <- paste0('DEgenes_C', cluster.ID, '_day_', day.ID)
  remaining <- c('remaining', 1, 'DEgenes')
  cvfit.dev.max <- rbind(cvfit.dev.max, remaining)
  
  
  # Predict clusters ysing cvfit and testing dataset
  predict.clusters <- predict(cvfit, 
                              newx = test, 
                              type = "class",
                              s = cvfit$lambda.min)
  
  # Return output
  return(list(prediction = predict.clusters, sub.cvfit.out = sub.cvfit.out, cvfit.dev.max = cvfit.dev.max, cvfit = cvfit))
  
}
```


```{r define_output_function}
comb <- function(x, ...) {
  lapply(seq_along(x),
    function(i) c(x[[i]], lapply(list(...), function(y) y[[i]])))
}
```



```{r define_ExtractResults_function}

ExtractResults <- function(training, prediction, cell.ids.cluster){

  # Compare to original clutering classes to check for accuracy ------------
  predict.index <- which(cell.ids.cluster[,1] %in% row.names(training))
  
  original.cluster <- cell.ids.cluster[predict.index,]
  original.cluster <- original.cluster[order(original.cluster[,2], decreasing = TRUE),]
  original.cluster <- as.data.frame(original.cluster)
  predict.clusters <- as.data.frame(prediction)
  predict.clusters$cellnames <- row.names(predict.clusters)
  compare <- merge(original.cluster, predict.clusters, by.x = 'V1', by.y = 'cellnames')
  
  # Change cluster IDs here
  cluster.select.predict <- subset(compare, (as.numeric(compare$cell.clusters) == cluster.ID &  compare$`1`==cluster.ID) | (compare$cell.clusters != cluster.ID &  compare$`1`!=cluster.ID))
  
  # Get accuracy
  accurate <- nrow(cluster.select.predict)
  inaccurate <- nrow(compare) - nrow(cluster.select.predict)
  list.acc.inacc <- list(accurate, inaccurate)
  
  # Return results ---------------------------------------------------------
  return(list(list.acc.inacc = list.acc.inacc, list.sig.genes = predict.marker$sub.cvfit.out, list.deviance = predict.marker$cvfit.dev.max, list.cvFit = predict.marker$cvfit)) 
}

```


```{r setup_cluster}
n_cores <- 2
cl <- makeCluster(n_cores, outfile = '') 
registerDoParallel(cl)
```


```{r define_iterations}
n <- 10
iter <- seq_len(n)  
```



```{r run_predictions}
#for DE genes
i <- 1
results <- foreach(i = iter, .combine = comb, .multicombine = TRUE, .init = list(list(), list(), list(), list(), list()), .packages = "glmnet") %dopar% {
  cat("\n##############################################################\n")
  cat('Iteration:', i, "\n")
  cat('----------', "\n")
  # Random sampling for training and testing datasets ----------------------
  data.train.test <- BuildTrainTest(clusterID.select = cluster.ID, 
                                    cluster.select = cluster.select, 
                                    clusterID.comp = 2, 
                                    cluster.compare = cluster.compare, 
                                    features = DE.gene.idx, 
                                    expression.data = expression.data)
  
  # Initialize variables
  test <- data.train.test$test
  training <- data.train.test$training
  response <- data.train.test$response
  
  cat('----------', "\n")
  cat('test    :', dim(test), "\n")
  cat('training:', dim(training), "\n")
  cat('response:', length(response),  "\n")
  cat("##############################################################\n")
  
  
 # Run lasso
  predict.marker <- FitRegModel(test, training, response, family = "binomial", alpha = 1)
  prediction <- predict.marker$prediction
  lasso <- ExtractResults(training = training, prediction = prediction, cell.ids.cluster = cell.ids.cluster)

  # Run elastic-net (alpha = 0.9)
  predict.marker <- FitRegModel(test, training, response, family = "binomial", alpha = 0.9)
  prediction <- predict.marker$prediction
  elastic.net.0.9 <- ExtractResults(training = training, prediction = prediction, cell.ids.cluster = cell.ids.cluster)
  
  # Run elastic-net (alpha = 0.5)
  predict.marker <- FitRegModel(test, training, response, family = "binomial", alpha = 0.5)
  prediction <- predict.marker$prediction
  elastic.net.0.5 <- ExtractResults(training = training, prediction = prediction, cell.ids.cluster = cell.ids.cluster)

  # Run elastic-net (alpha = 0.1)
  predict.marker <- FitRegModel(test, training, response, family = "binomial", alpha = 0.1)
  prediction <- predict.marker$prediction
  elastic.net.0.1 <- ExtractResults(training = training, prediction = prediction, cell.ids.cluster = cell.ids.cluster)

  # Run ridge
  predict.marker <- FitRegModel(test, training, response, family = "binomial", alpha = 0)
  prediction <- predict.marker$prediction
  ridge <- ExtractResults(training = training, prediction = prediction, cell.ids.cluster = cell.ids.cluster)
  
  i <- i  + 1
  return(list(lasso, elastic.net.0.9, elastic.net.0.5, elastic.net.0.1, ridge))

}
```


```{r sets_listnames}
names(results) <- c('lasso', 'elastic.net.0.9', 'elastic.net.0.5', 'elastic.net.0.1', 'ridge')
```


```{r stop_cluster}
stopCluster(cl)
```

```{r save_results}
saveRDS(results, file = paste0('All_List_c', cluster.ID, 'vs_Remaining_Day',day.ID,'_Testing.RDS'))
```


